---
title: "Homework #1"
author: "Camilo Baquero Jimenez - c.baquero10@uniandes.edu.co"
date: "7/2/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("zoo")
library("plyr")
library("MASS")
library("leaps")
library("dplyr")

import.csv <- function(filename){
  return(read.csv(filename,sep="," ,header=TRUE))
}

write.csv <- function(ob, filename){
  write.table(ob, filename, quote=FALSE, sep=",", row.names=FALSE)
}
```

```{r data, cache = TRUE, echo=FALSE}
##############################
#prepare data for PCA analysis
##############################
mydata <-import.csv('SP500_close_price.csv')
date <- as.Date(as.character(mydata[, 1]), format="%Y-%m-%d")
myzoo <- zoo(mydata[,-1], date )
myzoo <- na.locf(myzoo) #impute missing values
prices2returns <- function(x) 100*diff(log(x)) #function to covert from price to return
log.return.zoo <- prices2returns(myzoo)
log.return.data <- coredata(log.return.zoo) #data
log.return.date <- time(log.return.zoo) #date
##############################
#prepare industry sector data
sectors <- import.csv('SP500_ticker.csv')
```

## Problem 1

Apply Principal Component Analysis to stock market index data.

#### a) Fit a PCA model to log returns derived from stock price data:

```{r princomp}
pca.model <- princomp(log.return.data)
```

##### 1. Scree plot of the distribution of variance

```{r a1}
plot(pca.model)
```

##### 2. Cumulative percentage of variance retained

```{r a2}
pcaCPV <- function(x) {
    x.var <- x$sdev ^ 2
    x.pvar <- x.var/sum(x.var)
    plot(cumsum(x.pvar),xlab="Principal component",ylab="Cumulative percentage of variance",ylim=c(0,1),type='b')
}

pcaCPV(pca.model)
```

##### 3. Principal components retained

In order to capture at least 80% of the total variance in data, the top two PCA components must be retained.

```
summary(pca.model)

## Importance of components:
##                             Comp.1     Comp.2
## Standard deviation     105.9339978 49.1304919
## Proportion of Variance   0.6618857  0.1423687
## Cumulative Proportion    0.6618857  0.8042544
```

&nbsp;

##### 4. Magnitude of the estimated reconstruction error

If we only retain the top two of the PCA components, the magnitude of the estimated reconstruction error is:
&nbsp;
$$100\%*\left(1- \frac{\sum_{l=3}^{471} \lambda_l}{\sum_{i=1}^{471} \lambda_i} \right)=19.57456\%$$

&nbsp;
&nbsp;
&nbsp;

#### b) Analysis of principal components and weights:

##### 1. Time series of the 1st principal component

```{r b1}
principalcomponent <- zoo(pca.model$scores[, 1], log.return.date)
autoplot.zoo(principalcomponent)
```

##### 2. Weights from PCA principal two components

```{r b2, message=FALSE, warning=FALSE}
pca.weights <- pca.model$loadings[,1:2]
summary(pca.weights)
biplot(pca.model)
```

##### 3. Principal component 1 mean weight grouped by industry sector

```{r b3, echo=FALSE, warning=FALSE, message=FALSE}
ticker <- rownames(pca.weights)
comp1 <- pca.weights[,1]
comp2 <- pca.weights[,2]

w1 <- data.frame(ticker, comp1, comp2)
sectors <- data.frame(lapply(sectors, trimws))
weight1 <- inner_join(sectors, w1)

#out <- weight1 %>% group_by(sector) %>% summarize(comp1 = mean(comp1), comp2 = mean(comp2))
out1 <- weight1 %>% group_by(sector) %>% summarize(comp1 = mean(comp1))
plot(out1)

levels(weight1$sector)
```

##### 4. Principal component 2 mean weight grouped by industry sector

```{r b4, echo=FALSE, warning=FALSE, message=FALSE}

out2 <- weight1 %>% group_by(sector) %>% summarize(comp2 = mean(comp2))
plot(out2)
```

##### 5. New stock index using one principal component to track the overall market tendencies

Applications/RStudio.app/Contents/MacOS/pandoc/pandoc +RTS -K512m -RTS Tarea1.utf8.md --to html --from markdown+autolink_bare_uris+ascii_identifiers+tex_math_single_backslash --output Tarea1.html --smart --email-obfuscation none --self-contained --standalone --section-divs